---
title: "Toy example in Section 4"
header-includes:
   - \usepackage{bbm}
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

This notebook contains the code of the paper "Learning Physics between Digital Twins with Low-Fidelity Models and Physics-Informed Gaussian Processes". The models are fitted in rstan and the code is available in the folder "STAN/toy". 

#### Load packages

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message=FALSE,
  warning = FALSE,
  comment = '', 
  fig.width = 6, 
  fig.height = 3,
  fig.align = 'center'
)
```

```{r}
# uncomment to install
# install.packages("rstan")
# install.packages("ggplot2")
# install.packages("SAVE")
# install.packages("latex2exp")
# install.packages("ggpubr")

library(rstan)
library(ggplot2)
library(ggpubr)
library(SAVE) # package with the data
library(latex2exp)

rstan_options(auto_write = TRUE)
options(mc.cores = 3) # allocate 3 cores (for each model we run 3 chains in parallel)
```

#### Reality and modelling choice

\begin{align*}
    y^\mathcal{R}(x) &= 3.5\cdot \exp(-u\cdot x)+b+\varepsilon \quad  \text{ (the model we use to simulate data) }\\
    \eta(x,u) &= 5\cdot \exp(-u\cdot x) \quad \text{ (the misspesified model we use to fit the data) }
\end{align*}

```{r}
R = function(u,x,b) 3.5*exp(-u*x)+b
sd_noise = 0.3
```


```{r}
data("synthfield") # data from the rpackage SAVE
X_loc = unique(synthfield$x) 

# simulate data for different u val and add iid noise
u_val = seq(0.8,1.7,by=0.1)
dl=list()
set.seed(123)
offsets=runif(length(u_val),0.5,5)

xobs = c(unique(synthfield$x),unique(synthfield$x),unique(synthfield$x)); N=length(xobs); # create 3 replicates 
X_mat = matrix(NA, nrow = length(u_val), ncol = length(xobs))
set.seed(0)
dev=c(runif(length(u_val), 0.1,0.25)) 

for(i in 1:nrow(X_mat)){
  X_mat[i,] = xobs+dev[i] # create different input locations for each individual
}
for(i in seq_along(u_val)){
  set.seed(0)
  y=R(u_val[i], xobs, offsets[i])+rnorm(N,0,sd=sd_noise); # add i.i.d. N(0,0.3^2) noise
  dl[[i]] = list(x= X_mat[i,], y = y, N = N, x_pred=X_mat[i,], N_pred=N)
}
y = matrix(NA, nrow = length(u_val), ncol = length(xobs))
for(i in 1:nrow(y)) y[i,] = dl[[i]]$y

id = seq_along(u_val) # individual id
Ns = length(id) # total number of individuals
data_population = list(x= X_mat, y = y, N = N, Ns=Ns, id=id)
```

```{r echo=TRUE}
# create real data (noise free) for plotting 
X = seq(min(X_mat), max(X_mat), length.out = 50)
y_real = matrix(NA, nrow = Ns, ncol = length(X))
for(i in seq_along(u_val)){
  y_real[i,] = R(u_val[i],X,offsets[i]) # real data for plotting
}
id_new = paste0("ID = ", 1:Ns)
real_data = data.frame(x=rep(X,Ns), y = as.vector(t(y_real)), ID = rep(id_new, each=length(X)))
obs_data = data.frame(x=as.vector(t(X_mat)), y = as.vector(t(y)), ID = rep(id_new, each=ncol(X_mat)))

pl_obs_toy=ggplot()+
  geom_line(data=real_data, aes(x=x,y=y), color="dodgerblue3")+
  geom_point(data=obs_data, aes(x=x,y=y), color="black")+
  facet_wrap(ID~., nrow = 2)#+theme_bw()
pl_obs_toy
```

### Model 1 (no-without $\delta(x)$ in paper, Figure 4)

This is the misspecified model that does not account for model discrepancy (no-without delta in paper, Figure 3), and it is the following regression model
$$
y(x_m) = 5\cdot \exp(-u_m\cdot x_m) + \varepsilon, \text{ where }\varepsilon \sim N(0,\sigma^2).
$$
This is the following probabilistic model
$$
\mathbf{y} \sim N(5\cdot \exp(-u_m\cdot \mathbf{X}_m), \sigma^2I),
$$
where we assign priors to $u_m$ (same for each individual) and $\sigma$ (for more details see Appendix).

Stan code:
```{r}
# writeLines(readLines('STAN/toy/toy_nodelta.stan'))
```
We fit the model to each individual data set and we plot the trace for the last individual
```{r}
lu_no=list()
for(i in seq_along(u_val)){
  fit_no_without_delta = stan(
    file='STAN/toy/toy_nodelta.stan', # without delta
    data=dl[[i]],
    chains=3,
    iter=2*1000,
    seed=123
  )
  lu_no[[i]] = extract(fit_no_without_delta)$u
}
stan_trace(fit_no_without_delta, size=0.2)
```


### Model 2 (no-with $\delta(x)$ in paper, Figure 4)

Now we account for model discrepancy $\delta_m(x_m)\sim GP(0,K_\delta(x_m,x_m')),$ where we use the squared exponential kernel $K_\delta(x_m,x_m') = \alpha_m^2 \exp\left(-\frac{(x_m-x_m')^2}{2\rho_m^2}\right)$ and we have the following formulation
$$
y^R(\mathbf{x}) = \eta(\mathbf{x},\boldsymbol{\phi})+\delta(\mathbf{x})+\varepsilon, \text{ where } \varepsilon \sim N(0,\sigma^2).
$$
This is equivalent to
$$
y^\mathcal{R} \sim GP(5\cdot \exp(-u_m\cdot \mathbf{X}_m), K_\delta(\mathbf{X}_m,\mathbf{X}_m\mid \boldsymbol{\omega}_m)+\sigma^2I).
$$
Stan code:
```{r}
# writeLines(readLines('STAN/toy/toy_delta.stan'))
```

We fit this model to each individual data set separately

```{r}
lu=list()
for(i in seq_along(u_val)){
  fit_no_with_delta = stan(
    file='STAN/toy/toy_delta.stan', # with delta
    data=dl[[i]],
    chains=3,
    iter=2*1000,
    seed=123
  )
  lu[[i]] = extract(fit_no_with_delta)$u
}
stan_trace(fit_no_with_delta, size=0.2)
```


### Model 3 (yes/common $\delta(x)$ in paper, Figure 4)

We allow individuals to share information about the physical parameters $u_m, m=1,2,\ldots,10$ through a global level parameter as described in Section 3.2. The model assumes same discrepancy parameters for all individuals.

Stan code:
```{r}
# writeLines(readLines('STAN/toy/toy_common_delta.stan'))
```


```{r}
# shared u common delta model
fit_yes_common_delta = stan(file='STAN/toy/toy_common_delta.stan',
                            data=data_population,
                            chains=3,
                            iter=2*1000,
                            seed=123
)
names(fit_yes_common_delta)
stan_trace(fit_yes_common_delta, pars = "u", size=0.2)
ex_ycd=extract(fit_yes_common_delta)
```

### Model 4 (yes/shared $\delta(x)$ in paper, Figure 4)

We allow individuals to share information about both the physical parameters $u_m, m=1,2,\ldots,10$ and the discrepancy through a global level parameters for bothas described in Section 3.1. The model assumes same discrepancy parameters for all individuals.

Stan code:
```{r}
# writeLines(readLines('STAN/toy/toy_shared_delta.stan'))
```

```{r}
fit_yes_shared_delta = stan(file='STAN/toy/toy_shared_delta.stan',
                            data=data_population,
                            chains=3,
                            iter=2*1000,
                            seed=123
)
names(fit_yes_shared_delta)
stan_trace(fit_yes_shared_delta, pars = "u", size=0.2)
ex_ysd=extract(fit_yes_shared_delta)
```

#### Plot 95% CIs for all methods (Figure 3 in the paper) 

```{r echo=TRUE,fig.width = 8, fig.height = 3}
library(latex2exp)
nodelta_cis=m_nwd=matrix(NA,length(u_val),2)
fn_s = function(x) c(quantile(x, probs = c(0.025,0.975)))

for(i in seq_along(u_val)){
  nodelta_cis[i,] = fn_s(lu_no[[i]])
  m_nwd[i,] = fn_s(lu[[i]])
}

ysd_cis=data.frame(t(apply(ex_ysd$u,2,quantile,probs=c(0.025,0.975))))
yes_common_delta=data.frame(t(apply(ex_ycd$u,2,quantile,probs=c(0.025,0.975))))

m_nwd=data.frame(m_nwd)
colnames(m_nwd) = colnames(ysd_cis)
nodelta_cis=data.frame(nodelta_cis)
colnames(nodelta_cis) = colnames(ysd_cis)
m_nwd$u_true = u_val
m_nwd$sharing= "no-with delta"
nodelta_cis$u_true = u_val
nodelta_cis$sharing= "no-without delta"
yes_common_delta$u_true=u_val
yes_common_delta$sharing="yes/common delta"
ysd_cis$u_true=u_val
ysd_cis$sharing="yes/shared delta"

df_CIs = rbind(m_nwd,yes_common_delta,ysd_cis,nodelta_cis)
colnames(df_CIs)[1:2] = c("lower","upper")
df_CIs$id = rep(id,4)
lbl=c(latex2exp('no-without $\\delta(x)$'), 
                 latex2exp('no-with $\\delta(x)$'), 
                 latex2exp('yes/common $\\delta(x)$'), 
                 latex2exp('yes/shared $\\delta(x)$'))
toy_cis=ggplot(df_CIs,aes(x = as.factor(id), y = u_true, ymin = lower, ymax = upper, color=sharing))+ 
    geom_errorbar(width = 0.3, size=0.9) +
    theme_classic()+ 
    geom_point(size = 1.5, color="black")+
    ylab("u")+xlab("ID")+
    annotate("text", x=1.5, y=4.2, label= "95% CIs", size=6)+
    theme(
        legend.position="bottom",
        legend.key.size = unit(0.8, 'cm'), #change legend key size
        legend.key.height = unit(0.8, 'cm'), #change legend key height
        legend.key.width = unit(0.8, 'cm'), #change legend key width
        legend.title = element_text(size=12), #change legend title font size
        legend.text = element_text(size=12),
        axis.title.y = element_text(size = 16))+
    scale_color_manual(
      breaks=c('no-without delta', 'no-with delta', "yes/common delta", "yes/shared delta"),
      values=c("magenta","firebrick","dodgerblue4", "forestgreen"),
      labels = lbl)
toy_cis
# ggsave("Figs/toy_cis.pdf", plot = toy_cis, width = 20, height = 11, units = "cm")

```


```{r,fig.width = 8, fig.height = 2.5}
ID = paste0("ID = ",1:10)
df_CIs$ID = rep(ID,4)
rid=df_CIs$id%in%c(1,4,7,9)
toy_cis_new=ggplot()+ 
  geom_errorbar(data=df_CIs[rid,],aes(y = sharing, xmin = lower, xmax = upper, color=sharing), width = 0.3, size=0.9) +
  geom_vline(data=df_CIs[rid,],aes(xintercept=u_true),size = .5, color="black", linetype=2)+
  theme_classic()+ 
  facet_wrap(ID~., nrow = 1)+
  theme_bw()+
  theme(
    #strip.text.x = element_text(size = 15),
    axis.title.y=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank(),
    legend.position="bottom",
    legend.key.size = unit(0.8, 'cm'), #change legend key size
    legend.key.height = unit(0.8, 'cm'), #change legend key height
    legend.key.width = unit(0.8, 'cm'), #change legend key width
    legend.title = element_text(size=12), #change legend title font size
    legend.text = element_text(size=12),
    axis.title.x = element_text(size=12))+
  scale_color_manual(
    breaks=c('no-without delta', 'no-with delta', "yes/common delta", "yes/shared delta"),
    values=c("magenta","firebrick","dodgerblue4", "forestgreen"),
    labels = lbl)
toy_cis_new+coord_equal(ratio = 0.82)
# ggsave("Figs/toy_cis_new.pdf", plot = toy_cis_new+coord_equal(ratio = 0.82), width = 20, height = 8, units = "cm")
```

#### Uncertainty reduction

```{r}
df_red=df_CIs[df_CIs$sharing == "no-with delta" | df_CIs$sharing == "yes/common delta" | df_CIs$sharing == "yes/shared delta",]
df_red$dif = df_red$upper-df_red$lower
methd=c("no-with delta", "yes/common delta",  "yes/shared delta")
red_common = (df_red$dif[df_red$sharing == methd[1]] - df_red$dif[df_red$sharing == methd[2]])/ df_red$dif[df_red$sharing == methd[1]]
red_shared = (df_red$dif[df_red$sharing == methd[1]] - df_red$dif[df_red$sharing == methd[3]])/ df_red$dif[df_red$sharing == methd[1]]
mu_lu_fn = function(df) c(mu = mean(df), lower=min(df), upper=max(df))
# mu_lu_fn(red_common)
# mu_lu_fn(red_shared)
res = data.frame(common = as.vector(mu_lu_fn(red_common)), shared = as.vector(mu_lu_fn(red_shared)))
# rownames(res) = c("mean", "low", "upper")
# saveRDS(res, file = "Data/reduction_toy.rds")
```

#### Introduction Figure

```{r,fig.width = 7, fig.height = 4}
i=2
post_nnd = data.frame(u=lu_no[[i]],sharing="no discrepancy")
post_nwd = data.frame(u=lu[[i]],sharing="with discrepancy")
post_ysd = data.frame(u=ex_ysd$u[,i],sharing="our approach")
post_u = rbind(post_nnd, post_nwd, post_ysd)

pl_no_d=ggplot()+
  geom_violin(data=post_u[post_u$sharing == "no discrepancy",], aes(x=sharing, y=u), size=0.8, color="magenta")+
  geom_point(data = data.frame(u = u_val[i], sharing = "no discrepancy"), 
             aes(x = sharing, y = u, shape = "true"),size=2)+
  theme_bw()+
  theme(legend.position = "none",
        plot.title = element_text(face="bold"),
        axis.text = element_text(face="bold"),
        # panel.grid.major = element_blank(), 
        # panel.grid.minor = element_blank(),
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank()
        )+ ylim(0,5)+
  xlab("")+ ylab("u")+
  ggtitle("no discrepancy")

pl_with_d=ggplot()+
  geom_violin(data=post_u[post_u$sharing == "with discrepancy",], aes(x=sharing, y=u), size=1.2, color="firebrick")+
  geom_point(data = data.frame(u = u_val[i], sharing = "with discrepancy"), 
             aes(x = sharing, y = u, shape = "true"),size=2)+
  theme_bw()+
  theme(legend.position = "none",
        plot.title = element_text(face="bold"),
        axis.text = element_text(face="bold"),
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank()
        )+ ylim(0,5)+
  xlab("")+ ylab("")+
  ggtitle("with discrepancy")

pl_our=ggplot()+
  geom_violin(data=post_u[post_u$sharing == "our approach",], aes(x=sharing, y=u), size=1.2, color="dodgerblue4")+
  geom_point(data = data.frame(u = u_val[i], sharing = "our approach"), 
             aes(x = sharing, y = u, shape = "true"),size=2)+
  theme_bw()+
  theme(legend.position = "none",
        plot.title = element_text(face="bold"),
        axis.text = element_text(face="bold"),
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank()
        )+ ylim(0,5)+
  xlab("")+ ylab("")+
  ggtitle("our approach")

(pl_intro=ggarrange(pl_no_d, pl_with_d, pl_our, nrow=1))
# ggsave("Figs/intro2.pdf", plot = pl_intro, width = 18, height = 7, units = "cm")

```

### Session information

```{r}
sessionInfo()
```