---
title: "Cardiovascular model in Appendix E"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

This notebook contains the code of the paper "Learning Physics between Digital Twins with Low-Fidelity Models and Physics-Informed Gaussian Processes". The models are fitted in rstan and the code is available in the folder "STAN/WK2". 

#### Load packages

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message=FALSE,
  warning = FALSE,
  comment = '', 
  fig.width = 6, 
  fig.height = 6,
  fig.align = 'center'
)
```

```{r}
# uncomment to install
# install.packages("rstan")
# install.packages("ggplot2")
# install.packages("latex2exp")
library(rstan)
library(ggplot2)
library(latex2exp)

rstan_options(auto_write = TRUE)
options(mc.cores = 3) # allocate 3 cores (for each model we run 3 chains in parallel)
# Load flow data 
d = readRDS("Data/Inflow_time.rds")
```

#### Reality and modelling choice 

\begin{align}
  \mathcal{R}: \quad & \frac{d P(t)}{d t} + \frac{P(t)}{R_2C} = \frac{Q(t)}{C} \left (1 + \frac{ R_1}{R_2} \right ) + R_1 \frac{d Q(t)}{dt} \quad \text{ (the misspesified model we use to fit the data) }\text{ [WK3] }\\
  \eta: \quad  &  Q(t) = \frac{1}{R}P(t) + C \frac{dP(t)}{dt} \quad  \text{ (the model we use to simulate data) } \text{ [WK2] }
\end{align}

### Load posterior samples 

First, we load the posterior samples from the fitted models obtained using the 'toy_paper_code.Rmd' or in the experiments folder, 'WK2_appendix.R'.

```{r eval=TRUE}
# choose some reasonable physical parameter values
library(ggplot2)
dat = readRDS("Data/post_wk2.rds") # posterior data for all models
flow_time=readRDS("Data/Inflow_time.rds")
time=flow_time$time
post_nwd = dat$df_ycd
attach(dat)
attach(data)
lP_pr = lI_pr = list()
t1=Sys.time()
R_val=c(1,1.15, 1.3); C_val = c(1.1,0.95,1.25) 
RC=expand.grid(R_val,C_val) # create all possible combinations
Rtrue=RC[,1]; Ctrue=RC[,2]
nc=2
nP=12
nI=14
Ns=9
set.seed(123)
Zvec=sample(seq(0.02,0.1,by=0.01), Ns, replace = T)
nP_pred=50; nI_pred=50;
tP_pred = time[round(seq(1,length(time), length.out = nP_pred))]
tI_pred = time[round(seq(1,length(time), length.out = nI_pred))]
```


### Model 1 (no-without delta in paper, Figure 6)

This is the misspecified model that does not account for model discrepancy (no-without delta in paper, Figure 6). For more details on prediction equations see Appendix C.2.

Stan code:
```{r}
writeLines(readLines("STAN/WK2/WK2_nodelta_pred.stan"))
```

We predict for each individual $m=1,\ldots,9$

```{r}
#----------------------------------------------------------------
# no-without delta
#----------------------------------------------------------------
post_nnd=l_df_nnd
N_samples = nrow(post_nnd[[1]])
# no-without delta model
for (i in 1:Ns){
  data_pred = list(nP=nc*nP, nI=nc*nI, tP=tP[i,], tI=tI[i,], yP=yP[i,], yI=yI[i,]
                   , tP_pred = tP_pred, tI_pred=tI_pred, nP_pred=nP_pred, nI_pred=nI_pred
                   , alpha=post_nnd[[i]]$alpha, rho=post_nnd[[i]]$rho
                   , sigmaP=post_nnd[[i]]$sigmaP, sigmaI=post_nnd[[i]]$sigmaI
                   , R=post_nnd[[i]]$R, C=post_nnd[[i]]$C, N_samples=N_samples
  )
  
  pred = stan(file = "STAN/WK2/WK2_nodelta_pred.stan",
              data = data_pred,
              chains = 1, iter = 1, seed=123,
              algorithm = "Fixed_param")
  ex_pred=extract(pred)
  Pvec = ex_pred$y_P
  Ivec = ex_pred$y_I
  P_post=matrix(NA, ncol = 4, nrow = nP_pred)
  # calculate mean and 95% CIs
  for(n in 1:nP_pred){
    ind=(1 +(n-1)*N_samples):(n*N_samples)
    v = Pvec[ind]
    P_post[n,] = c(mean(v), quantile(v, probs = 0.025), quantile(v, probs = 0.975),tP_pred[n])
  }
  I_post=matrix(NA, ncol = 4, nrow = nI_pred)
  for(n in 1:nI_pred){
    ind=(1 +(n-1)*N_samples):(n*N_samples)
    v = Ivec[ind]
    I_post[n,] = c(mean(v), quantile(v, probs = 0.025), quantile(v, probs = 0.975),tI_pred[n])
  }
  
  lP_pr[[i]]=P_post
  lI_pr[[i]]=I_post
}
# prepare the data for plotting
ID = paste0(Zvec, " | ID = ", 1:Ns)
R1 = paste("expression(R[1] = )", Zvec)
P_nnd = data.frame(do.call(rbind, lP_pr))
P_nnd$ID = rep(ID, each = nP_pred)
P_nnd$t = rep(tP_pred, Ns)
colnames(P_nnd)[1:4] = c("mean", "lower", "upper", "time")
P_nnd$sharing = "no-without delta"

I_nnd = data.frame(do.call(rbind, lI_pr))
I_nnd$ID = rep(ID, each = nI_pred)
I_nnd$t = rep(tI_pred, Ns)
colnames(I_nnd)[1:4] = c("mean", "lower", "upper", "time")
I_nnd$sharing = "no-without delta"
```


### Model 2 (no-with delta in paper, Figure 6)

Now we account for model discrepancy $\delta_m(x_m)\sim GP(0,K_\delta(x_m,x_m')),$ where we use the squared exponential kernel $K_\delta(x_m,x_m') = \alpha_m^2 \exp\left(-\frac{(x_m-x_m')^2}{2\rho_m^2}\right).$ More details on the prediction equations are given in the Appendix C.2.

Stan code:
```{r}
writeLines(readLines("STAN/WK2/WK2_delta_pred.stan"))
```

We predict for each individual $m=1,\ldots,9$

```{r}
post_nwd=l_df_nwd
N_samples = nrow(post_nwd[[1]])
# no-with delta model
for (i in 1:Ns){
  data_pred = list(nP=nc*nP, nI=nc*nI, tP=tP[i,], tI=tI[i,], yP=yP[i,], yI=yI[i,]
                   , tP_pred = tP_pred, tI_pred=tI_pred, nP_pred=nP_pred, nI_pred=nI_pred
                   , alpha=post_nwd[[i]]$alpha, rho=post_nwd[[i]]$rho, alpha_d=post_nwd[[i]]$alpha_d
                   , rho_d=post_nwd[[i]]$rho_d, sigmaP=post_nwd[[i]]$sigmaP, sigmaI=post_nwd[[i]]$sigmaI
                   , R=post_nwd[[i]]$R, C=post_nwd[[i]]$C, N_samples=N_samples
  )
  
  pred = stan(file = "STAN/WK2/WK2_delta_pred.stan",
              data = data_pred,
              chains = 1, iter = 1, seed=123,
              algorithm = "Fixed_param")
  ex_pred=extract(pred)
  Pvec = ex_pred$y_P
  Ivec = ex_pred$y_I
  P_post=matrix(NA, ncol = 4, nrow = nP_pred)
  for(n in 1:nP_pred){
    ind=(1 +(n-1)*N_samples):(n*N_samples)
    v = Pvec[ind]
    P_post[n,] = c(mean(v), quantile(v, probs = 0.025), quantile(v, probs = 0.975),tP_pred[n])
  }
  I_post=matrix(NA, ncol = 4, nrow = nI_pred)
  for(n in 1:nI_pred){
    ind=(1 +(n-1)*N_samples):(n*N_samples)
    v = Ivec[ind]
    I_post[n,] = c(mean(v), quantile(v, probs = 0.025), quantile(v, probs = 0.975),tI_pred[n])
  }
  
  lP_pr[[i]]=P_post
  lI_pr[[i]]=I_post
}
# prepare data for plotting
# ID = paste0(Zvec, " | ID = ", 1:Ns)
P_nwd = data.frame(do.call(rbind, lP_pr))
P_nwd$ID = rep(ID, each = nP_pred)
P_nwd$t = rep(tP_pred, Ns)
colnames(P_nwd)[1:4] = c("mean", "lower", "upper", "time")
P_nwd$sharing = "no-with delta"

I_nwd = data.frame(do.call(rbind, lI_pr))
I_nwd$ID = rep(ID, each = nI_pred)
I_nwd$t = rep(tI_pred, Ns)
colnames(I_nwd)[1:4] = c("mean", "lower", "upper", "time")
I_nwd$sharing = "no-with delta"
```


### Model 3 (yes/common delta, Figure 6)

We allow individuals to share information about the physical parameters $u_m, m=1,2,\ldots,10$ through a global level parameter as described in Section 3.2. The model assumes same discrepancy parameters for all individuals.

We predict using the same equations as before, but now with the posterior distributions obtained by the fit of Model 3 (yes/common delta).

```{r}
post_ycd = list()

for(i in 1:Ns){
  v=c(paste0("alpha.",i), paste0("rho.",i), "alpha_d", "rho_d", "sigmaP", "sigmaI", paste0("R.",i), paste0("C.", i))
  df_temp=df_ycd[, v]
  colnames(df_temp)=c("alpha", "rho", "alpha_d", "rho_d", "sigmaP", "sigmaI", "R", "C")
  post_ycd[[i]]=df_temp
}

# yes/common delta model
for (i in 1:Ns){
  data_pred = list(nP=nc*nP, nI=nc*nI, tP=tP[i,], tI=tI[i,], yP=yP[i,], yI=yI[i,]
                   , tP_pred = tP_pred, tI_pred=tI_pred, nP_pred=nP_pred, nI_pred=nI_pred
                   , alpha=post_ycd[[i]]$alpha, rho=post_ycd[[i]]$rho, alpha_d=post_ycd[[i]]$alpha_d
                   , rho_d=post_ycd[[i]]$rho_d, sigmaP=post_ycd[[i]]$sigmaP, sigmaI=post_ycd[[i]]$sigmaI
                   , R=post_ycd[[i]]$R, C=post_ycd[[i]]$C, N_samples=N_samples
  )
  
  pred = stan(file = "STAN/WK2/WK2_delta_pred.stan",
              data = data_pred,
              chains = 1, iter = 1, seed=123,
              algorithm = "Fixed_param")
  ex_pred=extract(pred)
  Pvec = ex_pred$y_P
  Ivec = ex_pred$y_I
  P_post=matrix(NA, ncol = 4, nrow = nP_pred)
  for(n in 1:nP_pred){
    ind=(1 +(n-1)*N_samples):(n*N_samples)
    v = Pvec[ind]
    P_post[n,] = c(mean(v), quantile(v, probs = 0.025), quantile(v, probs = 0.975),tP_pred[n])
  }
  I_post=matrix(NA, ncol = 4, nrow = nI_pred)
  for(n in 1:nI_pred){
    ind=(1 +(n-1)*N_samples):(n*N_samples)
    v = Ivec[ind]
    I_post[n,] = c(mean(v), quantile(v, probs = 0.025), quantile(v, probs = 0.975),tI_pred[n])
  }
  
  lP_pr[[i]]=P_post
  lI_pr[[i]]=I_post
}

# ID = paste0(Zvec, "| ID = ", 1:Ns)
P_ycd = data.frame(do.call(rbind, lP_pr))
P_ycd$ID = rep(ID, each = nP_pred)
P_ycd$t = rep(tP_pred, Ns)
colnames(P_ycd)[1:4] = c("mean", "lower", "upper", "time")
P_ycd$sharing = "yes/common delta"

I_ycd = data.frame(do.call(rbind, lI_pr))
I_ycd$ID = rep(ID, each = nI_pred)
I_ycd$t = rep(tI_pred, Ns)
colnames(I_ycd)[1:4] = c("mean", "lower", "upper", "time")
I_ycd$sharing = "yes/common delta"
```

### Model 4 (yes/shared delta, Figure 6)

We allow individuals to share information about both the physical parameters $u_m, m=1,2,\ldots,9$ and the discrepancy through a global level parameters for both as described in Section 3.1. The model assumes same discrepancy parameters for all individuals.

We predict using the same equations as before, but now with the posterior distributions obtained by the fit of Model 4 (yes/shared delta).

```{r}
#-------------------------------------------------------
# yes/ shared delta
#----------------------------------------------------------------
post_ysd = list()

for(i in 1:Ns){
  v=c(paste0("alpha.",i), paste0("rho.",i), paste0("alpha_d.",i), paste0("rho_d.",i), "sigmaP", "sigmaI", paste0("R.",i), paste0("C.", i))
  df_temp=df_ysd[, v]
  colnames(df_temp)=c("alpha", "rho", "alpha_d", "rho_d", "sigmaP", "sigmaI", "R", "C")
  post_ysd[[i]]=df_temp
}

# yes/common delta model
for (i in 1:Ns){
  data_pred = list(nP=nc*nP, nI=nc*nI, tP=tP[i,], tI=tI[i,], yP=yP[i,], yI=yI[i,]
                   , tP_pred = tP_pred, tI_pred=tI_pred, nP_pred=nP_pred, nI_pred=nI_pred
                   , alpha=post_ysd[[i]]$alpha, rho=post_ysd[[i]]$rho, alpha_d=post_ysd[[i]]$alpha_d
                   , rho_d=post_ysd[[i]]$rho_d, sigmaP=post_ysd[[i]]$sigmaP, sigmaI=post_ysd[[i]]$sigmaI
                   , R=post_ysd[[i]]$R, C=post_ysd[[i]]$C, N_samples=N_samples
  )
  
  pred = stan(file = "STAN/WK2/WK2_delta_pred.stan",
              data = data_pred,
              chains = 1, iter = 1, seed=123,
              algorithm = "Fixed_param")
  ex_pred=extract(pred)
  Pvec = ex_pred$y_P
  Ivec = ex_pred$y_I
  P_post=matrix(NA, ncol = 4, nrow = nP_pred)
  for(n in 1:nP_pred){
    ind=(1 +(n-1)*N_samples):(n*N_samples)
    v = Pvec[ind]
    P_post[n,] = c(mean(v), quantile(v, probs = 0.025), quantile(v, probs = 0.975),tP_pred[n])
  }
  I_post=matrix(NA, ncol = 4, nrow = nI_pred)
  for(n in 1:nI_pred){
    ind=(1 +(n-1)*N_samples):(n*N_samples)
    v = Ivec[ind]
    I_post[n,] = c(mean(v), quantile(v, probs = 0.025), quantile(v, probs = 0.975),tI_pred[n])
  }
  
  lP_pr[[i]]=P_post
  lI_pr[[i]]=I_post
}
t2=Sys.time()

# ID = paste0(Zvec, "| ID = ", 1:Ns)
P_ysd = data.frame(do.call(rbind, lP_pr))
P_ysd$ID = rep(ID, each = nP_pred)
P_ysd$t = rep(tP_pred, Ns)
colnames(P_ysd)[1:4] = c("mean", "lower", "upper", "time")
P_ysd$sharing = "yes/shared delta"

I_ysd = data.frame(do.call(rbind, lI_pr))
I_ysd$ID = rep(ID, each = nI_pred)
I_ysd$t = rep(tI_pred, Ns)
colnames(I_ysd)[1:4] = c("mean", "lower", "upper", "time")
I_ysd$sharing = "yes/shared delta"
```

#### Plot pressure predictions (Figure 5 in Appendix)

```{r fig.width = 8, fig.height = 6}
P_pred = rbind(P_nwd, P_ysd, P_ycd, P_nnd)
df_true = dat$P_true
P_true = data.frame(P=as.vector(dat$P_true), t = rep(time,Ns), ID = rep(ID, each = length(time)))

P_obs = data.frame(P=as.vector(t(yP)), t = as.vector(t(tP)), ID = rep(ID, each = nc*nP))

# New facet label names 
appender = function(string) {
  TeX(paste("$R_1= $", string))
}

pl_WK_pred_P=ggplot()+
  geom_point(data=P_obs, aes(x=t, y=P, shape="observed"))+
  geom_ribbon(data = P_pred, aes(x=t,ymin=lower,ymax=upper, fill=sharing),alpha=0.3)+# alpha=0.38
  geom_line(data=P_pred, aes(x=t, y=mean, color= sharing), size=0.7)+
  geom_line(data=P_true, aes(x=t, y=P, linetype="true"), size=0.6)+
  facet_wrap(~ID, nrow = 3,labeller = as_labeller(appender, default = label_parsed))+
  theme_bw()+
  xlab("time (sec)")+ylab("pressure (mmHg)")+
  theme(legend.position = "bottom", legend.title = element_blank())+
  scale_color_manual(
    breaks=c('no-without delta','no-with delta', "yes/common delta", "yes/shared delta", "true"),
    values=c("magenta","firebrick","dodgerblue4", "forestgreen", "black")
  )+
  scale_linetype_manual(breaks=c('no-without delta','no-with delta', "yes/common delta", "yes/shared delta", "true"),
                        values=c(1,1,1, 1, 2)
  )+
  scale_fill_manual(
    breaks=c('no-without delta','no-with delta', "yes/common delta", "yes/shared delta"),
    values=c("magenta","firebrick","dodgerblue4", "forestgreen")
  )

pl_WK_pred_P
# ggsave("Figs/WK_pred_P.pdf", plot = pl_WK_pred_P, width = 20, height = 15, units = "cm")
```


#### Plot flow predictions (Figure 6 in Appendix)

```{r fig.width = 8, fig.height = 6}
I_pred = rbind(I_nwd, I_ysd, I_ycd, I_nnd)
df_true_I = dat$I_true
I_true = data.frame(I=as.vector(flow_time$inflow), t = rep(time,Ns), ID = rep(ID, each = length(time)))

I_obs = data.frame(I=as.vector(t(yI)), t = as.vector(t(tI)), ID = rep(ID, each = nc*nI))
unique(I_pred$sharing)


pl_WK_pred_I=ggplot()+
  geom_point(data=I_obs, aes(x=t, y=I, shape="observed"))+
  geom_ribbon(data = I_pred, aes(x=t,ymin=lower,ymax=upper, fill=sharing),alpha=0.3)+# alpha=0.38
  geom_line(data=I_pred, aes(x=t, y=mean, color= sharing), size=0.7)+
  geom_line(data=I_true, aes(x=t, y=I, linetype="true"), size=0.6)+
  facet_wrap(~ID, nrow = 3,labeller = as_labeller(appender, default = label_parsed))+
  theme_bw()+
  xlab("time (sec)")+ylab("flow (ml/min)")+
  theme(legend.position = "bottom", legend.title = element_blank())+
  scale_color_manual(
    breaks=c('no-without delta','no-with delta', "yes/common delta", "yes/shared delta", "true"),
    values=c("magenta","firebrick","dodgerblue4", "forestgreen", "black")
  )+
  scale_linetype_manual(breaks=c('no-without delta','no-with delta', "yes/common delta", "yes/shared delta", "true"),
                        values=c(1,1,1, 1, 2)
  )+
  scale_fill_manual(
    breaks=c('no-without delta','no-with delta', "yes/common delta", "yes/shared delta"),
    values=c("magenta","firebrick","dodgerblue4", "forestgreen")
  )

pl_WK_pred_I
# ggsave("Figs/WK_pred_I.pdf", plot = pl_WK_pred_I, width = 20, height = 15, units = "cm")
```

#### Plot posteriors of R for all individuals (Figure 3 in Appendix)

```{r fig.width = 8, fig.height = 6}
R_nnd = R_nwd = R_ycd = R_ysd = C_nnd = C_nwd = C_ycd = C_ysd = c()

for(i in 1:Ns){
  R_temp = post_nnd[[i]]$R
  R_nnd = c(R_nnd, R_temp)
  R_temp = post_nwd[[i]]$R
  R_nwd = c(R_nwd, R_temp)
  R_temp = post_ycd[[i]]$R
  R_ycd = c(R_ycd, R_temp)
  R_temp = post_ysd[[i]]$R
  R_ysd = c(R_ysd, R_temp)
  C_temp = post_nnd[[i]]$C
  C_nnd = c(C_nnd, C_temp)
  C_temp = post_nwd[[i]]$C
  C_nwd = c(C_nwd, C_temp)
  C_temp = post_ysd[[i]]$C
  C_ycd = c(C_ycd, C_temp)
  C_temp = post_ycd[[i]]$C
  C_ysd = c(C_ysd, C_temp)
}
id_post = rep(as.factor(1:Ns), each = N_samples)
shr = c('no-without delta','no-with delta', "yes/common delta", "yes/shared delta")
Rpost=data.frame(R = c(R_nnd, R_nwd, R_ycd, R_ysd), ID = rep(id_post,4), sharing = rep(rep(shr, each = Ns*N_samples),4))
Cpost=data.frame(C = c(C_nnd, C_nwd, C_ycd, C_ysd), ID = rep(id_post,4), sharing = rep(rep(shr, each = Ns*N_samples),4))

# R post
(pl_R=ggplot() + 
    geom_violin(data = Rpost, aes(x = ID, y=R, color = sharing), size=0.7)+
    theme_bw()+ 
    theme(axis.title.y = element_text(size = rel(1.5)), legend.position = "none")+
    geom_point(data = data.frame(R = Rtrue, ID = as.factor(1:Ns)), aes(x = ID, y = R, shape = "true"),size=1)+
    facet_wrap(sharing~.)+
    xlab("ID")+
    scale_color_manual(
      breaks=c('no-without delta', 'no-with delta', "yes/common delta", "yes/shared delta"),
      values=c("magenta","firebrick","dodgerblue4", "forestgreen"))
)
# ggsave("Figs/post_WK_R.pdf", plot = pl_R, width = 18, height = 12, units = "cm")
```


#### Plot posteriors of C for all individuals (Figure 3 in Appendix)

```{r fig.width = 8, fig.height = 6}
# C post
(pl_C=ggplot() + 
  geom_violin(data = Cpost, aes(x = ID, y=C, color = sharing),size=0.7)+
  theme_bw()+ 
  theme(axis.title.y = element_text(size = rel(1.5)), legend.position = "none")+
  geom_point(data = data.frame(C = Ctrue, ID = as.factor(1:Ns)), aes(x = ID, y = C, shape = "true"),size=1)+
  facet_wrap(sharing~.)+
  xlab("ID")+
  scale_color_manual(
    breaks=c('no-without delta', 'no-with delta', "yes/common delta", "yes/shared delta"),
    values=c("magenta","firebrick","dodgerblue4", "forestgreen")))
# ggsave("Figs/post_WK_C.pdf", plot = pl_C, width = 18, height = 12, units = "cm")
```


### Session information

```{r}
sessionInfo()
```