---
title: "Cardiovascular model in Section 5"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

This notebook contains the code of the paper "Learning Physics between Digital Twins with Low-Fidelity Models and Physics-Informed Gaussian Processes". The models are fitted in rstan and the code is available in the folder "STAN/WK2". 

#### Load packages

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message=FALSE,
  warning = FALSE,
  comment = '', 
  fig.width = 6, 
  fig.height = 6,
  fig.align = 'center'
)
```

```{r}
# uncomment to install
# install.packages("rstan")
# install.packages("ggplot2")
# install.packages("SAVE")
library(rstan)
library(ggplot2)

rstan_options(auto_write = TRUE)
options(mc.cores = 3) # allocate 3 cores (for each model we run 3 chains in parallel)

# Numerical simulator of the WK3 model
source("WK_numerical_simulators/WK2and3_sim_fn.R")
# Load flow data 
d = readRDS("Data/Inflow_time.rds")
```

#### Reality and modelling choice 

\begin{align}
  \mathcal{R}: \quad & \frac{d P(t)}{d t} + \frac{P(t)}{R_2C} = \frac{Q(t)}{C} \left (1 + \frac{ R_1}{R_2} \right ) + R_1 \frac{d Q(t)}{dt} \quad \text{ (the misspesified model we use to fit the data) }\text{ [WK3] }\\
  \eta: \quad  &  Q(t) = \frac{1}{R}P(t) + C \frac{dP(t)}{dt} \quad  \text{ (the model we use to simulate data) } \text{ [WK2] }
\end{align}



```{r eval=TRUE}
# choose some reasonable physical parameter values
R_val=c(1,1.15, 1.3); C_val = c(1.1,0.95,1.25) 
RC=expand.grid(R_val,C_val) # create all possible combinations
Rtrue=RC[,1]; Ctrue=RC[,2]
Ns=length(Rtrue) # number of individuals
flow = d$inflow*0.95 # flow data
time = d$time # corresponding time
nP=12 # number of pressure data
nI=14 # number of inflow data
nc=2  # number of cardiac cycles
nflow = length(flow)
post_nnd=post_nwd=l_df_nnd=l_df_nwd=list()

set.seed(123)
Zvec=sample(seq(0.02,0.1,by=0.01), length(Rtrue), replace = T)
yP=tP=matrix(NA,nrow = Ns, ncol = nP*nc)
yI=tI=matrix(NA,nrow = Ns, ncol = nI*nc)
P_true = matrix(NA, nrow = nflow, ncol = Ns)
t1=Sys.time()
for(i in 1:Ns){
  # 1. simulate WK3 data (R=R_2, Z=R_1)
  Psim=WK3_simulate(flow = flow, time = time, R = Rtrue[i], C = Ctrue[i], Z=Zvec[i]) # simulate WK3 data for a given flow Q(t)
  P_true[,i] = Psim
  # 2. choose pressure and inflow indices
  indP = round(seq(1, nflow, length.out = nP)); indI = round(seq(1, nflow, length.out = nI))
  yP_real = Psim[indP]; yI_real = flow[indI] # noise free fimulated pressure and flow
  # 3. Add noise
  set.seed(123)
  Pnoise = rnorm(nP*nc, 0, 4) # sample pressure noise from N(0, 4^2)
  Inoise =rnorm(nI*nc, 0, 10) # sample flow noise from N(0,10^2)
  yP_real = rep(yP_real,nc) # create 2 replicates (2 cardiac cycles/heart beats) 
  yI_real = rep(yI_real,nc) # create 2 replicates (2 cardiac cycles/heart beats)
  # 4. store individual data in the population matrices
  yP[i,]= yP_real + Pnoise # add noise
  yI[i,]= yI_real + Inoise # add noise
  tP[i,] = time[indP] # corresponding time (synchronized for the two cycles)
  tI[i,] = time[indI] # corresponding time (synchronized for the two cycles)
}
id=1:Ns
data_population = list(nP=nc*nP, nI=nc*nI, tP=tP, tI=tI, yP=yP, yI=yI,id=id, Ns=Ns)
```



```{r}
ID = paste0("ID = ", 1:Ns)
df_Ptrue = data.frame(pressure = as.vector(P_true), time = rep(time,Ns), ID = rep(ID, each=nflow))
df_Pobs = data.frame(pressure = as.vector(t(yP)), time=as.vector(t(tP)), ID = rep(ID, each=nP))

ggplot()+
  geom_line(data=df_Ptrue, aes(x=time,y=pressure), color="red")+
  geom_point(data=df_Pobs, aes(x=time,y=pressure), color="black", size=0.5)+
  facet_wrap(ID~., nrow = 3)+theme_bw()+xlab("time (sec)")+ ylab("pressure (mmHg)")

```

### Model 1 (no-without delta in paper, Figure 6)

This is the misspecified model that does not account for model discrepancy (no-without delta in paper, Figure 6). For more details on the physics-informed Gaussian process prior see Appendix E.1.

Stan code:
```{r}
writeLines(readLines('STAN/WK2/WK2_nodelta.stan'))
```

We fit the model to each individual data set and we plot the trace for the last individual
```{r}
#-------------------------------------------------
### Model 1 (no-without delta in paper, Figure 6)
# WK2 PI prior / no delta (magenta model)
for(i in 1:Ns){
  data_ind = list(nP=nc*nP, nI=nc*nI, tP=tP[i,], tI=tI[i,], yP=yP[i,], yI=yI[i,])
  fit_nnd= stan(file= 'STAN/WK2/WK2_nodelta.stan', 
                    data=data_ind,
                    chains=3,
                    iter=1000,
                    seed=123
  )
  post_nnd[[i]]=extract(fit_nnd)
  l_df_nnd[[i]]=data.frame(extract(fit_nnd))
}
stan_trace(fit_nnd, size=0.2)
```


### Model 2 (no-with delta in paper, Figure 6)

Now we account for model discrepancy $\delta_m(x_m)\sim GP(0,K_\delta(x_m,x_m')),$ where we use the squared exponential kernel $K_\delta(x_m,x_m') = \alpha_m^2 \exp\left(-\frac{(x_m-x_m')^2}{2\rho_m^2}\right).$ More details about the model are given in the Appendix E.1.

Stan code:
```{r}
writeLines(readLines('STAN/WK2/WK2_delta.stan'))
```

We fit this model to each individual data set separately

```{r}
# WK2 PI prior / with pressure delta (red model)
for(i in 1:Ns){
  data_ind = list(nP=nc*nP, nI=nc*nI, tP=tP[i,], tI=tI[i,], yP=yP[i,], yI=yI[i,])
  fit_nwd= stan(file= 'STAN/WK2/WK2_delta.stan', #'STAN_WK/WK_ind_delta.stan',
                          data=data_ind,
                          chains=3,
                          iter=1000,
                          seed=123
  )
  post_nwd[[i]]=extract(fit_nwd)
  l_df_nwd[[i]]=data.frame(extract(fit_nwd))
}
stan_trace(fit_nwd, size=0.2, pars = c("R", "C"))
```


### Model 3 (yes/common delta, Figure 6)

We allow individuals to share information about the physical parameters $u_m, m=1,2,\ldots,10$ through a global level parameter as described in Section 3.2. The model assumes same discrepancy parameters for all individuals.

Stan code:
```{r}
writeLines(readLines('STAN/WK2/WK2_common_delta.stan'))
```

```{r}
#-------------------------------------------------
# shared R,C common delta (blue model)
fit_yes_common_delta = stan(file='STAN/WK2/WK2_common_delta.stan',
                            data=data_population,
                            chains=3,
                            iter=1000,
                            seed=123
)
names(fit_yes_common_delta)
stan_trace(fit_yes_common_delta, pars = c("R","C"), size=0.2)
ex_ycd=extract(fit_yes_common_delta)
df_ycd =data.frame(ex_ycd)
```

### Model 4 (yes/shared delta, Figure 6)

We allow individuals to share information about both the physical parameters $u_m, m=1,2,\ldots,9$ and the discrepancy through a global level parameters for both as described in Section 3.1. The model assumes same discrepancy parameters for all individuals.

Stan code:
```{r}
writeLines(readLines('STAN/WK2/WK2_common_delta.stan'))
```

Model fit:

```{r}
#-------------------------------------------------
# shared R,C shared delta (green model)
fit_ysd = stan(file='STAN/WK2/WK2_shared_delta.stan',
                            data=data_population,
                            chains=3,
                            iter=1000,
                            seed=123
)
names(fit_ysd)
stan_trace(fit_ysd, pars = c("R", "C"), size=0.2)
ex_ysd=extract(fit_ysd)
df_ysd=data.frame(ex_ysd)
```

#### Plot 95% CIs for all methods (Figure 6 in the paper) 

```{r echo=FALSE,fig.width = 6, fig.height = 4}
l_post_wk2 = list(l_df_nnd=l_df_nnd, l_df_nwd=l_df_nwd, df_ycd=df_ycd, df_ysd=df_ysd, data=data_population, P_true=P_true)
# saveRDS(l_post_wk2, "Data/post_wk2.rds")
# common delta
post_common = data.frame(extract(fit_yes_common_delta))
Rpost_common=post_common[,paste0("R.",1:length(Rtrue))]
Cpost_common=post_common[,paste0("C.",1:length(Rtrue))]
# shared delta
post_shared = data.frame(extract(fit_ysd))
Rpost_shared=post_shared[,paste0("R.",1:length(Rtrue))]
Cpost_shared=post_shared[,paste0("C.",1:length(Rtrue))]
# indepedent 
Rpost_ind = Cpost_ind =Rpost_ind_delta =Cpost_ind_delta = matrix(NA,nrow = nrow(Rpost_common), ncol = ncol(Rpost_common))

for(i in 1:length(Rtrue)){
  Rpost_ind[,i]=post_nnd[[i]]$R
  Cpost_ind[,i]=post_nnd[[i]]$C
  Rpost_ind_delta[,i]=post_nwd[[i]]$R
  Cpost_ind_delta[,i]=post_nwd[[i]]$C
}

# 95% CIs
R_par_pool_common_CI=data.frame(t(apply(Rpost_common,2,quantile, probs=c(0.025, 0.975))))
C_par_pool_common_CI=data.frame(t(apply(Cpost_common,2,quantile, probs=c(0.025, 0.975))))
R_par_pool_shared_CI=data.frame(t(apply(Rpost_shared,2,quantile, probs=c(0.025, 0.975))))
C_par_pool_shared_CI=data.frame(t(apply(Cpost_shared,2,quantile, probs=c(0.025, 0.975))))
R_sep_CI=data.frame(t(apply(Rpost_ind,2,quantile, probs=c(0.025, 0.975))))
C_sep_CI=data.frame(t(apply(Cpost_ind,2,quantile, probs=c(0.025, 0.975))))
R_sep_CI=data.frame(t(apply(Rpost_ind,2,quantile, probs=c(0.025, 0.975))))
C_sep_CI=data.frame(t(apply(Cpost_ind,2,quantile, probs=c(0.025, 0.975))))
R_sep_delta_CI=data.frame(t(apply(Rpost_ind_delta,2,quantile, probs=c(0.025, 0.975))))
C_sep_delta_CI=data.frame(t(apply(Cpost_ind_delta,2,quantile, probs=c(0.025, 0.975))))


colnames(R_par_pool_common_CI) = c("low", "upper")
colnames(C_par_pool_common_CI) = c("low", "upper")
colnames(R_par_pool_shared_CI) = c("low", "upper")
colnames(C_par_pool_shared_CI) = c("low", "upper")
colnames(R_sep_CI) = c("low", "upper")
colnames(C_sep_CI) = c("low", "upper")
colnames(R_sep_delta_CI) = c("low", "upper")
colnames(C_sep_delta_CI) = c("low", "upper")


R_par_pool_common_CI$par="R"
R_par_pool_common_CI$id=id
R_par_pool_shared_CI$par="R"
R_par_pool_shared_CI$id=id
R_par_pool_common_CI$sharing="yes/common delta"
R_par_pool_common_CI$true = Rtrue+Zvec
R_par_pool_shared_CI$sharing="yes/shared delta"
R_par_pool_shared_CI$true = Rtrue+Zvec
C_par_pool_common_CI$par="C"
C_par_pool_common_CI$id=id
C_par_pool_common_CI$true=Ctrue
C_par_pool_common_CI$sharing="yes/common delta"
C_par_pool_shared_CI$par="C"
C_par_pool_shared_CI$id=id
C_par_pool_shared_CI$true=Ctrue
C_par_pool_shared_CI$sharing="yes/shared delta"

R_sep_CI$par="R"
R_sep_CI$id=id
R_sep_CI$true=Rtrue+Zvec
R_sep_CI$sharing="no-without delta"
C_sep_CI$par="C"
C_sep_CI$id=id
C_sep_CI$true=Ctrue
C_sep_CI$sharing="no-without delta"

R_sep_delta_CI$par="R"
R_sep_delta_CI$id=id
R_sep_delta_CI$true=Rtrue+Zvec
R_sep_delta_CI$sharing="no-with delta"
C_sep_delta_CI$par="C"
C_sep_delta_CI$id=id
C_sep_delta_CI$true=Ctrue
C_sep_delta_CI$sharing="no-with delta"

df_CIs = rbind(R_sep_delta_CI,C_sep_delta_CI,
               R_par_pool_shared_CI,C_par_pool_shared_CI,
               R_par_pool_common_CI,C_par_pool_common_CI,
               R_sep_CI,C_sep_CI)

saveRDS(df_CIs, file = "WK_fig_data.rds")
readRDS("WK_fig_data.rds")
(pl_cis=ggplot(df_CIs)+
    geom_errorbar(aes(ymin=low, ymax=upper, x=as.factor(id), color=sharing),size=1)+
    geom_point(aes(x=as.factor(id),y=true))+
    facet_wrap(par~.)+theme_bw()+ylim(0.5,3)+xlab("ID")+ylab("95% CIs")+
    #scale_color_discrete(breaks=c('no without delta', 'no with delta', "yes/common delta", "yes/shared delta"))+
    scale_color_manual(
      breaks=c('no-without delta', 'no-with delta', "yes/common delta", "yes/shared delta"),
      values=c("magenta","firebrick","dodgerblue4", "forestgreen")))
# ggsave("Figs/wk_cis.pdf", plot = pl_cis, width = 20, height = 8, units = "cm")
```


### Session information

```{r}
sessionInfo()
```